Fusion Moves for Correlation Clustering

### 2014-11-13 paper/introduction-new.tex
update introduction/notation

In computer vision applications, unsupervised image segmentation algorithms usually start with an over-segmentation
into superpixels (superregions), which are then clustered into ``perceptually meaningful''
regions by correlation clustering on sparse graphs.
Such an approach has been shown to yield
state-of-the-art results on the Berkeley Segmentation Database
\cite{andres_2011_iccv,yarkony_2012_eccv,alush_2013_simbad}.

@@

% \paragraph{Contribution.} The evident usefulness of correlation clustering and its clean and compact formulation in terms of an optimization problem, together with its unfortunate NP-hardness, are an invitation to develop fast approximate solvers. The basic idea of the move making algorithm proposed here is to maintain, at all times, a best current partitioning; and to iteratively improve it (as we show, monotonously) by considering diverse and cheaply generated proposal partitionings. Any two nodes that are in one cluster in both the current and the proposal partitioning are contracted. Edges to contracted nodes are equally combined and reweighted, and the correlation clustering problem is then solved on this reduced graph. We offer a polyhedral characterization of this strategy, and evaluate it in conjunction with two versatile proposal generators.   We conduct experiments on a broad range of data sets ranging from 2D and 3D segmentation problems to clustering in signed social networks. The results suggest that this simple move making algorithm is the fastest technique known today, reaching close to globally optimal solutions in a tenth or a hundredth of the time required by the most efficient exact solvers. 
\vspace{0.1cm}
\noindent \textbf{Contribution.}
In this work we present some novel approaches that are designed for large scale correlation clustering problems.
First, we define a novel energy based agglomerative clustering algorithm that monotonically increase the energy.
With this at hand we show how to improve the anytime performance of Cut, Clue \& Cut.
Second, we introduce cluster-fusion moves, which extend the original fusion moves~\cite{Lempitsky-2010} 
used in supervised segmentation to the unsupervised case and give a polyhedral interpretation of this algorithm.
We propose two versatile proposal generators, and evaluate the proposed methods on existing and new benchmark problems.
Experiments show that we can improve the computation time by one to two magnitudes without worsening the segmentation 
quality significantly.

\vspace{0.1cm}
\noindent \textbf{Related Work.}
A natural approach is to solve the integer linear programming problem directly. 
To this end, efficient separation procedures have been found~\cite{kappes_2011_emmcvpr,kappes_2013_arxiv} that allow to iteratively augment the set of constraints until a valid partitioning is found. 
Alternatively, it is possible to relax the integrality constraints of the integer linear programming formulation~\cite{kappes_2013_arxiv}. 
Such an outer relaxation can be iteratively tightened. However, intermediate solutions are fractional and therefore rounding is required to obtain a valid partitioning.
For the latter approach column generating methods exits, that works best on planar \cite{yarkony_2012_eccv} graphs. %or almost planar \cite{yarkony-andres-2013} graphs.

Another line of work uses move making algorithms 
to optimize correlation clustering~\cite{bansal_2004_ml,beier_2014_cvpr,Kernighan-1970}.
Starting with an initial segmentation, auxiliary max-cut problems are approximately solved,
such that the segmentation is strictly improved.
As shown by Beier \etal~\cite{beier_2014_cvpr} only CGC~\cite{beier_2014_cvpr} 
can deal with large scale problems, but can also suffer on very large auxiliary problems.

@@

For energy minimization problems fusion moves have become increasingly popular~\cite{Lempitsky-2010,kappes_2014_ws}.
For many large scale computer vision applications fusion moves lead to good approximations
with state of the art any time performance~\cite{kappes_2014_ws}.
Due to the ambiguity of a node-labeling, classical fusion moves~\cite{Lempitsky-2010} can not be applied directly for correlation clustering.
We will show how to overcome this point later.

\vspace{0.1cm}
\noindent \textbf{Outline:} 
In Sec.~\ref{sec:problem_formulation} we will give a 
detailed problem definition where we introduce 
the correlation clustering objective.
In Sec.~\ref{sec:cc_fm} we describe our proposed
method and show the properties of the algorithm.
In Sec.~\ref{sec:exp} we show an evaluation
of the proposed methods methods and state of the art
and conclude in Sec.~\ref{sec:future} and \ref{sec:conclusion}.
%-------------------------------------------------------------------------

\section{Notation and Problem Formulation}\label{sec:problem_formulation}
Let $G=(V,E, w)$ be a weighted graph of nodes $V$ and edges $E$.
%
The function $w : E \rightarrow \mathbb{R}$ assigns a weight to each edge.
We will use $w_e$ as a shorthand for $w(e)$.
A positive weight expresses the desire that two adjacent nodes should
be merged, whereas a negative weight indicates
that these nodes should be separated into two different regions.
%
%A \emph{subgraph} $G_A = \{A, E_A, w\}$ consists
%of nodes $A \subseteq V$ and edges $E_A := E\cap (A\times A)$.
%
A segmentation of the graph $G$ can by either given by an 
node labeling $l \in \mathbb{N}^{|V|}$
or an edge labeling $y \in\{0,1\}^{|E|}$, \cf Fig.~\ref{fig:notation}.
An edge labeling is only consistent if it contains no dangling edges~\cite{kappes_2013_arxiv}.
We denote the set of all consistent edge labelings by $P(G)\subset\{0,1\}^{|E|}$.
The convex hull of this set is known as the \emph{multicut polytope} $MC(G) = \textrm{conv}(P(G))$.

@@

\input{inputs/fig_notation.tex}

>>>>

In computer vision applications, unsupervised image segmentation algorithms usually start with an over-segmentation
into superpixels (superregions), which are then clustered into ``perceptually meaningful''
regions by correlation clustering on sparse graphs.
Such an approach has been shown to yield
state-of-the-art results on the Berkeley Segmentation Database
\cite{andres_2011_iccv,Kim-2011,yarkony_2012_eccv,alush_2013_simbad}.

@@

% \paragraph{Contribution.} The evident usefulness of correlation clustering and its clean and compact formulation in terms of an optimization problem, together with its unfortunate NP-hardness, are an invitation to develop fast approximate solvers. The basic idea of the move making algorithm proposed here is to maintain, at all times, a best current partitioning; and to iteratively improve it (as we show, monotonously) by considering diverse and cheaply generated proposal partitionings. Any two nodes that are in one cluster in both the current and the proposal partitioning are contracted. Edges to contracted nodes are equally combined and reweighted, and the correlation clustering problem is then solved on this reduced graph. We offer a polyhedral characterization of this strategy, and evaluate it in conjunction with two versatile proposal generators.   We conduct experiments on a broad range of data sets ranging from 2D and 3D segmentation problems to clustering in signed social networks. The results suggest that this simple move making algorithm is the fastest technique known today, reaching close to globally optimal solutions in a tenth or a hundredth of the time required by the most efficient exact solvers. 
\vspace{0.1cm}
\noindent \textbf{Contribution.}
In this work we present some novel approaches that are designed for large scale correlation clustering problems.
First, we define a novel energy based agglomerative clustering algorithm that monotonically increase the energy.
With this at hand we show how to improve the anytime performance of Cut, Clue \& Cut~\cite{beier_2014_cvpr}.
%
Second, we improve the anytime performance of polyhedral multicut methods~\cite{kappes_2013_arxiv} by more efficient separation procedures.
%
Third, we introduce cluster-fusion moves, which extend the original fusion moves~\cite{Lempitsky-2010} 
used in supervised segmentation to the unsupervised case and give a polyhedral interpretation of this algorithm.
We propose two versatile proposal generators, and evaluate the proposed methods on existing and new benchmark problems.
Experiments show that we can improve the computation time by one to two magnitudes without worsening the segmentation 
quality significantly.

\vspace{0.1cm}
\noindent \textbf{Related Work.}
A natural approach is to solve the integer linear program (ILP) directly. 
To this end, efficient separation procedures have been found~\cite{kappes_2011_emmcvpr,kappes_2013_arxiv} that allow to iteratively augment the set of constraints until a valid partitioning is found. 
Alternatively, it is possible to relax the integrality constraints of the ILP formulation~\cite{kappes_2013_arxiv}. 
Such an outer relaxation can be iteratively tightened. However, intermediate solutions are fractional and therefore rounding is required to obtain a valid partitioning.
For the latter approach column generating methods exits, that works best on planar \cite{yarkony_2012_eccv} graphs. %or almost planar \cite{yarkony-andres-2013} graphs.

Another line of work uses move making algorithms 
to optimize correlation clustering~\cite{bansal_2004_ml,beier_2014_cvpr,Kernighan-1970}.
Starting with an initial segmentation, auxiliary max-cut problems are (approximately) solved,
such that the segmentation is strictly improved.
As shown by Beier \etal~\cite{beier_2014_cvpr} only Cut, Clue \& Cut (CGC)~\cite{beier_2014_cvpr} 
can deal with large scale problems, but can also suffer on very large auxiliary problems.

@@

For energy minimization problems fusion moves have become increasingly popular~\cite{Lempitsky-2010,kappes_2014_ws}.
For many large scale computer vision applications fusion moves lead to good approximations
with state of the art any time performance~\cite{kappes_2014_ws}.
Due to the ambiguity of a node-labeling, classical fusion moves~\cite{Lempitsky-2010} can not be applied directly for correlation clustering.
We will show how to overcome this point in Sec.~\ref{sec:cc_fm}.

\vspace{0.1cm}
\noindent \textbf{Outline:} 
In Sec.~\ref{sec:problem_formulation} we will give a 
detailed problem definition where we introduce 
the correlation clustering objective.
Followed by a description of energy based hierarchical clustering in Sec.~\ref{sec:ehc} and
our proposed correlation clustering fusion moves in Sec.~\ref{sec:cc_fm}.
We evaluate the proposed methods methods in Sec.~\ref{sec:exp} and conclude in Sec.~\ref{sec:future} and \ref{sec:conclusion}.

%-------------------------------------------------------------------------

\input{inputs/fig_notation.tex}
\section{Notation and Problem Formulation}\label{sec:problem_formulation}
Let $G=(V,E, w)$ be a weighted graph of nodes $V$ and edges $E$.
%
The function $w : E \rightarrow \mathbb{R}$ assigns a weight to each edge.
We will use $w_e$ as a shorthand for $w(e)$.
A positive weight expresses the desire that two adjacent nodes should
be merged, whereas a negative weight indicates
that these nodes should be separated into two different regions.
%
%A \emph{subgraph} $G_A = \{A, E_A, w\}$ consists
%of nodes $A \subseteq V$ and edges $E_A := E\cap (A\times A)$.
%
%\input{inputs/fig_notation.tex}
A segmentation of the graph $G$ can by either given by an 
node labeling $l \in \mathbb{N}^{|V|}$
or an edge labeling $y \in\{0,1\}^{|E|}$, \cf Fig.~\ref{fig:notation}.
An edge labeling is only consistent if it contains no dangling edges~\cite{kappes_2013_arxiv}.
We denote the set of all consistent edge labelings by $P(G)\subset\{0,1\}^{|E|}$.
The convex hull of this set is known as the \emph{multicut polytope} $MC(G) = \textrm{conv}(P(G))$.
%
By $l(y)$ we denote a node labeling for a segmentation given by $y$.

@@

%\input{inputs/fig_notation.tex}
